---
title: "Homework 1"
author: "Charandeep Grewal 301190855, James Hopkins 301139318, Alisha McGrath 301242597"
date: "September 12, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Question 1 (Chapter 2, #1, 8 marks)

(a) With a large sample size and a small number of predictors we should use a flexible method. There is a smaller chance of overfitting when the dataset is large.

(b) With small data sets and a large number of predictors, the chance for overfitting the training data will be high. Therefore Training MSE will be small, but the test MSE may be large

(c) 
 Flexible models may be necessary to detect non-linear trends

(d)



## Question 2 (Chapter 2, #2, 6 marks)

(a) 
   Regression, Inference, n = 500, p = 4
(b)
   Classification, Prediction, n = 20, p = 13
(c)
   Regression, Prediction, n = 52, p =3

```{r}
library(ISLR)
data(Auto)
head(Auto)
```
9a)

    Qualitative: cylinders, year, origin, name
    Quantitative: mpg, displacement, horsepower, weight, acceleration
9b)
```{r}
sapply(Auto[, (1:7)], range)
```

9c) 
```{r}
sapply(Auto[, (1:7)], mean)
sapply(Auto[, (1:7)], sd)
```

9d)
```{r}
Auto1 <- Auto[-10:-85, (1:7)]
sapply(Auto1[, (1:7)], range)
sapply(Auto[, (1:7)], mean)
sapply(Auto[, (1:7)], sd)
```

9e)
```{r}
library(ggplot2)
plot(Auto$cylinders, Auto$mpg,xlab="Cylinders",ylab="Mpg")
plot(Auto$weight, Auto$mpg,xlab="Weight",ylab="Mpg")
plot(Auto$horsepower, Auto$mpg,xlab="Horsepower",ylab="Mpg")
plot(Auto$year, Auto$mpg,xlab="Year",ylab="Mpg")
plot(Auto$acceleration, Auto$mpg,xlab="Acceleration",ylab="Mpg")
plot(Auto$displacement, Auto$mpg,xlab="Displacement",ylab="Mpg")
```