---
title: "Homework 3"
author: "Charandeep Grewal 301190855, James Hopkins 301139318, Alisha McGrath 301242597"
date: '2017-11-01'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Question 1 (Chapter 6, #8, parts (a)-(e), 10 marks)


(a) (1 mark)
```{r}
set.seed(94)
X = rnorm(100)
eps = rnorm(100)
```
(b) (1 mark)
```{r}
beta0 = 2
beta1 = -3
beta2 = 5
beta3 = 7
Y = beta0 + beta1 * X + beta2 * X^2 + beta3 * X^3 + eps
```

(c) (3 marks)
```{r}
library(leaps)
data.full = data.frame("y" = Y, "x" = X)
mod.full = regsubsets(y~poly(x, 10, raw=T), data=data.full, nvmax=10)
mod.summary = summary(mod.full)
which.min(mod.summary$cp)
which.min(mod.summary$bic)
which.max(mod.summary$adjr2)
```
For Cp, BIC, and adjusted R^2, the $3$ variable model is selected. 
```{r}
plot(mod.summary$cp, xlab="Subset Size", ylab="Cp",  type="l")
points(3, mod.summary$cp[3], col = "blue", lwd=7)
plot(mod.summary$bic, xlab="Subset Size", ylab="BIC", type="l")
points(3, mod.summary$bic[3], col="blue", lwd=7)
plot(mod.summary$adjr2, xlab="Subset Size", ylab="Adjusted R2", type="l")
points(3, mod.summary$adjr2[3], col="blue", lwd=7)
```
```{r}
coefficients(mod.full, id=3)
```
(d) (2 marks) 
```{r}
forward.model = regsubsets(y~poly(x, 10, raw=T), data=data.full, nvmax=10, method="forward")
back.model = regsubsets(y~poly(x, 10, raw=T), data=data.full, nvmax=10, method="backward")
fwd.summary = summary(forward.model)
bwd.summary = summary(back.model)
which.min(fwd.summary$cp)
which.min(bwd.summary$cp)
which.min(fwd.summary$bic)
which.min(bwd.summary$bic)
which.max(fwd.summary$adjr2)
which.max(bwd.summary$adjr2)
```
```{r}
par(mfrow=c(3, 2))
plot(fwd.summary$cp, xlab="Subset Size", ylab="Forward Cp", type="l")
points(3, fwd.summary$cp[3], col="blue", lwd=7)
plot(bwd.summary$cp, xlab="Subset Size", ylab="Backward Cp", type="l")
points(3, bwd.summary$cp[3], col="blue", lwd=7)
plot(fwd.summary$bic, xlab="Subset Size", ylab="Forward BIC", type="l")
points(3, fwd.summary$bic[3], col="blue", lwd=7)
plot(bwd.summary$bic, xlab="Subset Size", ylab="Backward BIC", type="l")
points(3, bwd.summary$bic[3], col="blue", lwd=7)
plot(fwd.summary$adjr2, xlab="Subset Size", ylab="Forward Adjusted R2", type="l")
points(3, fwd.summary$adjr2[3], col="blue", lwd=7)
plot(bwd.summary$adjr2, xlab="Subset Size", ylab="Backward Adjusted R2", type="l")
points(4, bwd.summary$adjr2[4], col="blue", lwd=7)
```
```{r}
coefficients(forward.model, id=3)
coefficients(back.model, id=3)
coefficients(back.model, id=4)
```
The coefficients are very similar to (c). Backward selection model with $4$ variables chooses to use $X^8$ for its 4th variable, however the varaible seems negligible with such a low coefficient.

(e) (3 marks)
```{r}
library(glmnet)
Xmat = model.matrix(y~poly(x, 10, raw=T), data=data.full)[, -1]
lambdas <- 10^{seq(from=-2,to=5,length=100)}
cv.lafit <- cv.glmnet(Xmat,Y,alpha=1,lambda=lambdas) 
plot(cv.lafit)
la.best.lam <- cv.lafit$lambda.1se
la.best.lam
la.best <- glmnet(Xmat,Y,alpha=1,lambda=la.best.lam)
coef(la.best)
```
The lasso model with the lambda chosen has $4$ variables rather than $3$ varaibles obtained from the majority of the previous models from parts (c) and (d). The lasso model chooses $X^5$ as its 4th variable instead of $X^8$ like the backward selection model with $4$ varaibles. However, $X^8$ has a small coefficient that could be negligible depending on context of our model. Also, the other $3$ variables chosen are the same as all the previous models with similar coefficients. Lastly, the lasso model shows that some of the varaibles have a coefficient of 0 and are not useful to be included in our model.   


## Question 2 (Ch6, #9, 12 marks)

(a) (0 marks)
To make everyone's results comparable, please
select your test set with the following.

```{r}
library(ISLR)
data(College)
dim(College) # 777 rows, use 111 as test
set.seed(1)
testset <- sample(1:777,size=111)
College.test <- College[testset,]
College.train <- College[-testset,]
```

(b) (2 marks)

(c) (2 marks)

(d) (2 marks)

(e) (2 marks)

(f) (2 marks)

(g) (2 marks)


## Question 3 (Ch7, #6, 8 marks)

(a) (5 marks)

(b) (3 marks)
